---
title: "Data Source Jam"
subtitle: "Project: CATE"
date: 2018-07-24  
output:
  word_document: default
  pdf_document: default
  html_document:
  code_folding: hide
  df_print: paged
---

# Data Exploration Loop 1
Data Exploration Topics

* Summary statistics (incl. mean, median, variations, ...)
* Five numbers summary by boxplots (numerical data)
* Distribution by bar charts and histograms
* Categories reduction
* Derived variables
* Missing values and outliers
* Subgroups, patterns and trends
* Covariations (relationships between pairs of attributes)
    - distribution of a continuous variable broken down by a categorical variable
    - between categorical variables
    - between two continuous variables
* Correlations (numerical variables)

```{r, echo = FALSE}
# ====================== Libraries ================
library(readxl)
library(tidyverse)
library(hexbin)
# ====================== Parameters ==============
rel_fpath  <- "./data/"    # relative directory where original data is stored
filename ='A5E01283425_30062016_25072016' # file name of the original data 
endung ='.csv'            # file extension of the original data
rel_fpath_output <- "./data/" # relative directory where outputs are stored
outputfile <- "./output/data_exploration.pdf"  # writes plots and clean data frame to outputfile
outputcsv  <- "./output/clean_4_modelling_with_R.csv"

Read_Fresh_from_File <- F  # Set to TRUE for first run to fill var rawframe
                           # clean dat frame, set to FALSE for initial data understanding
options(scipen=999)        # suppress exponential form of print output format
printpdf <- FALSE          # if printout in PDF desired
shortnames <- TRUE         # if T change to short variable names is desired
scatterplot <- FALSE       # it T do scatterplot
basic_tranform <- T        # do basic feature transformations
# ====================== Sources ===============
source(file = "Functions_data_exploration.R")
```
#2. D a t a   U n d e r s t a n d i n g
  
##2.1 Collect Initial Data

Routines:  
Read from data sources from "rel_fpath"!  
Read description and configuration file from "rel_fpath"!  
Save data as RDS file format 

```{r, echo = FALSE}
# ====================== R e a d   d a t a ============== alternative
library(writexl)
if (Read_Fresh_from_File)  {
  print("reading file.... ")
  # --- read excel ----
  if (endung == ".xlsx")   {
     rawframe <- read_xlsx (path = paste0(rel_fpath, filename, endung ), guess_max=100000)
     warnings()
  } else
  {  #--- read other formats ---
  library(readr)
  rawframe <- read_delim(file = paste0(rel_fpath, filename, endung), 
    ";", escape_double = FALSE, col_types = cols(DTDATUM = col_datetime(format = "%d.%m.%Y %H:%M:%S"), 
        DTDATUM_1 = col_datetime(format = "%d.%m.%Y %H:%M:%S"), 
        DTINSERT = col_datetime(format = "%d.%m.%Y %H:%M:%S"), 
        DTU = col_double(), SERGEBNIS = col_factor(levels = c("FAIL", 
            "PASS"))), locale = locale(decimal_mark = ","), 
    trim_ws = TRUE)
  }

  # initial exclusion of not relevant variables according to Data Description entry in column Clean Data Frame
  rawframe <- rawframe[ ,datadesc$`Clean Data Frame`!="exclude"]  

  saveRDS(rawframe, file = paste(rel_fpath_output , filename, ".rData"))
  print("...Data saved in RData file")
} else # Read_Fresh_from_File gleich FALSE
  rawframe <- readRDS(file = paste(rel_fpath_output , filename, ".rData"))

#--- create / read data control file ---
if (file.exists(path = paste0(rel_fpath, filename, '_data_control.xlsx')))
  data_control_file <-  paste0(rel_fpath, filename, '_data_control.xlsx') else
    data_control_file <- paste0(rel_fpath, 'data_control_template.xlsx')
datadesc <- read_datadesc(data_control_file)
  
# write control file if not present
if (!file.exists(path = paste0(rel_fpath, filename, '_data_control.xlsx')))  {
  datadesc[1:length(names(rawframe)), 1] <-  names(rawframe)
  writexl::write_xlsx(datadesc, path = paste0(rel_fpath, filename, '_data_control.xlsx'))
}

```

##2.2 Describe Data   

##2.3 Explore Data and Visualization

Routines:  
Create sub dataframes of numeric and non numeric variables  
Do basic data transformation and sorting  
```{r, echo = FALSE}
#sorting - eases further exploration
rawframe <- rawframe %>% arrange(DTDATUM, NORDER)

#basic transformations - datenspezifisch!
if (basic_tranform) {
  #df_num$`Total Customer Price (EUR)` <- log(df_num$`Total Customer Price (EUR)`)
  rawframe <- rawframe[rawframe$SEINHEIT != 'OE' ,]

#create a sub dataframe of numeric varialbes
df_num <- as.data.frame (rawframe [ , sapply(rawframe,is.numeric)])
print (paste("Dimension of dataframe of numeric variables (columns)", dim(df_num)[2]))

#create a sub dataframe of non numeric varialbes
df_not_num <- as.data.frame (rawframe [ , !sapply(rawframe,is.numeric)])
print (paste("Dimension of dataframe of not-numeric variables (columns)", dim(df_not_num)[2]))
}
```

### 2.3.1 Boxplots and Histograms of all numerical variables

This section shows boxplots, histograms and the summary statistics of all numerical variables as well as related comments and descriptions which are extracted from the data control file.

```{r, echo = FALSE}
if (printpdf) pdf(outputfile)
if (scatterplot) plot (df_num) # scatter plot
plot_num (df_num, datadesc)
if (printpdf) dev.off()    # reset device to screen plot 
```

Explore a single numerical variable - data specific - and detect patterns:
* peaks at 10x iterm
* counts
* 3 most common values
```{r, echo = FALSE}
# mini feature engineering CATE
above_thresh_df <- df_num$DMW[(df_num$DMW>df_num$DTO)] - df_num$DTO[(df_num$DMW>df_num$DTO)] %>% as.data.frame( )
names(above_thresh_df) <- "above_thresh"
below_thresh_df <- df_num$DTU[(df_num$DMW<df_num$DTU)] - df_num$DMW[(df_num$DMW<df_num$DTU)] %>% as.data.frame( )
names(below_thresh_df) <- "below_thresh"

# Limitiation of counts on y axis as well as outliers on x axis
ggplot(data = above_thresh_df) +
  geom_histogram(mapping = aes(x = above_thresh), binwidth = 50000, na.rm = TRUE) +
  coord_cartesian(xlim = c(0, 11000000), ylim = c(0, 100)) +
  labs(x = "above_thresh")

ggplot(data = above_thresh_df) +
  geom_histogram(mapping = aes(x = above_thresh), binwidth = 1000, na.rm = TRUE) +
  coord_cartesian(xlim = c(0, 22000), ylim = c(0, 10)) +
  labs(x = "above_thresh")

# use "group by"" and "summarize" allways together
c_items <- above_thresh_df %>%
  group_by(above_thresh) %>%
  summarise(count = n()) %>%
  print()

#print ("3 most common value")
c_items <- (sort(table(above_thresh_df$above_thresh), decreasing = TRUE)) %>%
     as.data.frame()
c_items$perc <- c_items$Freq/sum(c_items$Freq)*100
print(c_items[1:3,])

# Limitiation of counts on y axis as well as outliers on x axis
ggplot(data = below_thresh_df) +
  geom_histogram(mapping = aes(x = below_thresh), binwidth = 5000, na.rm = TRUE) +
  coord_cartesian(xlim = c(0, 300000), ylim = c(0, 100)) +
  labs(x = "below_thresh")

# Limitiation of counts on y axis as well as outliers on x axis
ggplot(data = below_thresh_df) +
  geom_histogram(mapping = aes(x = below_thresh), binwidth = 10, na.rm = TRUE) +
  coord_cartesian(xlim = c(0, 500), ylim = c(0, 100)) +
  labs(x = "below_thresh")

# use "group by"" and "summarize" allways together
c_items <- below_thresh_df %>%
  group_by(below_thresh) %>%
  summarise(count = n()) %>%
  print()

#print ("3 most common value")
c_items <- (sort(table(below_thresh_df$below_thresh), decreasing = TRUE)) %>%
     as.data.frame()
c_items$perc <- c_items$Freq/sum(c_items$Freq)*100
print(c_items[1:3,])

```

NOCH SIEMENS EM SPEZIFISCH:
Explore a single numerical variable, here "Probability of Order", dependent on categorical variable, here "Status"  
Order it by median of "Probability of Order"

Findings: ...

Problem bei Facetierung von Histogrammen: Countwerte dann inkonsistent wenn Spalte über ..$.. adressiert wird!
https://drsimonj.svbtle.com/plotting-background-data-for-groups-with-ggplot2
 
```{r, echo = FALSE}
# rawframe used as varialbes from df_num and df_not_num requested
ggplot(data = rawframe) +
        geom_boxplot(mapping = aes(x = reorder(rawframe$Status, rawframe$`Probability of Order`, FUN = median),
                                   y = rawframe$`Probability of Order`), na.rm = TRUE) + 
        labs(x = names(rawframe$`Probability of Order`), y = "")

ggplot(data = rawframe, mapping = aes(x = rawframe$`Probability of Order`)) +
         geom_freqpoly(mapping = aes(color = Status))
ggplot(data = rawframe, mapping = aes(x = rawframe$`Probability of Order`, y = ..density..)) +
         geom_freqpoly(mapping = aes(color = Status))
# subset
t <- filter (rawframe, rawframe$Status == "Offer won")
ggplot(data = t, mapping = aes(x = t$`Probability of Order`)) +
         geom_freqpoly(mapping = aes(color = Status))
ggplot(data = t, mapping = aes(x = t$`Probability of Order`)) +
         geom_histogram(binwidth = 1) +
         facet_wrap(~ Status, ncol = 2)

ggplot(data = rawframe, mapping = aes(x = `Probability of Order`)) +  # facet_wrap funktioniert nur wenn kein 'rawframe$' davor
         geom_histogram(binwidth = 1) +
         facet_wrap(~ Status, ncol = 2, scales = "free_y")
ggplot(data = rawframe, mapping = aes(x = `Probability of Order`, y = ..density..)) +
         geom_histogram(binwidth = 1) +
         facet_wrap(~ Status, ncol = 2)

```

### 2.3.2 Exploration of all non numerical variables

```{r, echo = FALSE}
summary_non_num (df_not_num, datadesc)
```

## 2.4 Verify data quality

### 2.4.1 check missing values

```{r}
j <- 1 #j is rawframe index, i is datadesc index
for(i in 1:nrow(datadesc)){  
  if (datadesc$`Clean Data Frame`[i] !='exclude') {
    cat(paste(names(rawframe[j]), "\t  # of NA:", colSums(is.na(rawframe[j])), 
              "  in %", round(colSums(is.na(rawframe[j])/nrow(rawframe)), 3)*100), "\n")
    datadesc$`% of na values`[i] <- round(colSums(is.na(rawframe[j]))/nrow(rawframe), 3)*100
    j <- j+1
  }  
}    
#write.csv2(datadesc$`% of na values`, file = "na_%.csv")
writexl::write_xlsx(datadesc, path = paste0(rel_fpath, filename, '_data_control.xlsx'))

```


### 2.4.2 check collinearity


```{r}
# produce correlation matrix
check_collinearity (df_num)
```

## 2.5 Pivot Section Board view - DF MC specific

```{r}

count_messw <- rawframe %>% group_by(SID) %>% summarize(count=n(), timediff=sum(DTINSERT-DTDATUM>100000))
count_messw_pass <- rawframe %>% filter(SERGEBNIS=='PASS') %>% group_by(SID) %>% summarize(count_pass=n(),
#                                                                                           time_pass_insert=list(as.character(unique(DTINSERT))),
                                                                                           time_pass=list(as.character(unique(DTDATUM))))
count_messw_fail <- rawframe %>% filter(SERGEBNIS=='FAIL') %>% group_by(SID) %>% summarize(count_fail=n(),
#                                                                                           time_fail_insert=list(as.character(unique(DTINSERT))),
                                                                                           time_fail=list(as.character(unique(DTDATUM))),
                                                                                           limit_viol=sum((DMW>DTO)|(DMW<DTU)))
count_messw <- count_messw %>% full_join (count_messw_pass) %>% full_join (count_messw_fail)

```

## 2.6 Combine Q-Daten

```{r}
q_daten <- read_delim("data/Q_Daten_FBG_ A5E01283425_04.07.2018.csv", 
    ";", escape_double = FALSE, col_types = cols(BEFUND_STATUS1 = col_skip(), 
        BEFUND_STATUS2 = col_skip(), PRUEFUNG = col_skip()), trim_ws = TRUE,
    n_max=16)  # skip questions and double entries

count_messw <- count_messw %>% full_join (q_daten, by = c("SID" = "FID")) 
```

```{r}
print("...save board data frame")
outputcsv <- "./output/board.csv"
write.table(count_messw,       # write resulting data into CSV file with tabstops
            file=outputcsv,
            sep=",", quote=FALSE, row.names=FALSE, dec=".") 
```

## 2.7 Trend failures - DF MC specific

```{r}

messw_fail_trend <- rawframe %>% filter(SERGEBNIS=='FAIL') %>% group_by(SFILENAME) %>% summarize(count_fail=n(),
                                                                                           time_fail=unique(DTDATUM))

ggplot(data = messw_fail_trend, mapping = aes(x = time_fail)) +
         geom_histogram(binwidth = 10000)

hist(messw_fail_trend$time_fail, 100)

```

## 2.8 Pivot Section Component view - DF MC specific

```{r}

board_view <- rawframe %>% group_by(SMMNR) %>% summarize(count=n(), limit_viol=sum((DMW>DTO)|(DMW<DTU)))

spec_bauteil <- rawframe %>% filter(SMMNR=='10.40.235_R701_235') %>% arrange(DTINSERT)
plot(spec_bauteil$DTINSERT, spec_bauteil$DMW, type ="l")

```

## 3.1 Transformation of variables

```{r}
datadesc <- read_datadesc(data_control_file)
df_clean <- transformation(rawframe, datadesc)

#create new sub dataframes for numeric and non numeric variables
df_num <- as.data.frame (df_clean [ , sapply(df_clean,is.numeric)])
df_not_num <- as.data.frame (df_clean [ , !sapply(df_clean,is.numeric)])
```

## 3.2 Reduction of variables, visualization and saving of clean data

NOT RELEVANT in DF MC case - already done above
```{r}
datadesc <- read_datadesc(data_control_file)
# exclude not relevant variables according to Data Description entry in column Clean Data Frame
df_clean <- rawframe[ ,datadesc$`Clean Data Frame`!="exclude"]  

#create new sub dataframes for numeric and non numeric variables
df_num <- as.data.frame (df_clean [ , sapply(df_clean,is.numeric)])
df_not_num <- as.data.frame (df_clean [ , !sapply(df_clean,is.numeric)])

```

```{r}
# =================================================
# again perform data exploration as above in chapter 2
                           # Boxplots and Histograms for numerical variables
if (printpdf) pdf(outputfile)
if (scatterplot) plot (df_num) # scatter plot
plot_num (df_num, datadesc)
if (printpdf) dev.off()    # reset device to screen plot 

summary_non_num (df_not_num, datadesc)    # Exploration of non numerical variables
```

```{r}
# specific line for DF MC
df_clean <- rawframe
print("...save clean data frame")
if (shortnames==T) outputcsv <- "./output/clean_4_modeling_with_factors.csv"
write.table(df_clean,       # write resulting data into CSV file with tabstops
            file=outputcsv,
            sep=",", quote=FALSE, row.names=FALSE, dec=".") 
```

## 3.3 Create dummy variables, visualization and saving of clean data

```{r}
datadesc <- read_datadesc(data_control_file)
df_clean <- create_dummy(df_clean, datadesc)

#create new sub dataframes for numeric and non numeric variables
df_num <- as.data.frame (df_clean [ , sapply(df_clean,is.numeric)])
df_not_num <- as.data.frame (df_clean [ , !sapply(df_clean,is.numeric)])
```

```{r}
# =================================================
# Save clean data frame with dummy variables
# =================================================
print("...save clean data frame")
if (shortnames==T) outputcsv <- "./output/clean_4_modeling_with_dummies.csv"
write.table(df_clean,       # write resulting data into CSV file with tabstops
            file=outputcsv, 
            sep=",", quote=FALSE, row.names=FALSE, dec=".") 
```

